```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This homework is about syntax and basic plotting functions in R, and is **due by the end of Feb 7**. 

You can use the RMarkDown file with these homework instructions as a template for your work. 
Make sure to replace "Your Name" and "your_github_username" in the YAML with your name and Github username.

**Submission instructions**: Create a folder named `ds202_hw2`, and name the RMarkDown file including your solutions `hw2.Rmd` under this folder. For submission, create a GitHub repository named `ds202_hw2` under your account, and push both `hw2.Rmd` and the knitted `hw2.html` to GitHub before the deadline. I will assume you use the same GitHub username as for your HW1 submission. The latest version of your homework appearing on GitHub before the deadline is graded. *It is crucial to follow the exact folder structure and file names*, so that your homework can be reproduced and graded by an automated script. Some questions will require the following answer format.

## Example question and answer
0\. What is the smallest positive odd integer? Storage it in a variable named `a0` and print it.

**Example answer**:
In the RMarkDown file, type in
<pre><code>```{r}
a0 <- 1
a0
```</code></pre>
The knitted html file will show:
```{r}
a0 <- 1
a0
```

## R commands
1. Today's temperature has a high of 24 and a low of 15 degree fahrenheit.  Find the corresponding degress in celsius and storage it in a (single) vector named `cel`. Print the variable.
```{r}
cel = c(-4.44444, -9.44444)
cel
```
2. Create a vector of length 200. The odd entries should be 3 and even entries should be 1. Storage the created in a variable named `vec200` and print.
```{r}
vec200 = rep(c(3,1), times=100)
vec200
```
3. How many perfect squares are there between 1 and 2001? Storage the answer in a  variable named `numPerfect` and print. You may want to look at `?round`.
```{r}
numPerfect = 0
for (i in seq(1, 2001)) {
  if (sqrt(i) %% 1 == 0) {
      numPerfect = numPerfect + 1
  }
}
numPerfect
```

## Cars and mileage
For this part we will use the `mpg` data set, you can access the data by first installing (using `install.packages('ggplot2')`) and then loading the package `ggplot2` (using `library(ggplot2)`) into your R session. The `mpg` data includes information about fuel economy of popular car models
for the years 1999 and 2008 as collected by the US Environmental Protection Agency. An answer should include a short paragraph of text for the explanation to a (human) reader, together with appropriate R code, outputs, and graphs.

Most exercises are from the ggplot2 book section 2.2.1. You might find it helpful for working on the homework to read through this section. *However, all questions can be solved by using just the basic R graphics functions.*

```{r}
library(ggplot2)
```

1. What are the 3 most fuel-efficient models in terms of highway mileage? Storage the data frame (with 3 rows) containing all information of these three cars in a variable named `top3` and print.

The following creates a dataframe sorted, which is created by sorting descending by the mpg$hwy column. The top 3 results are then the first 3 rows in the sorted dataframe.
```{r}
sorted = mpg[order(-mpg$hwy) , ]
top3 = sorted[1:3, ]
top3
```

2. How many different compact models are included in this dataset? Storage the answer in  a variable named `numCompact` and print.

For this problem I get the number of rows where the column mpg$class is equal to the string 'compact' and print it.
```{r}
numCompact = length(which(mpg$class == 'compact'))
numCompact
```

3. Create a scatterplot of `hwy` and `cty` and describe the relationship. Why are there so few points visible? Explain your findings.

This graph has a positive correlation. I believe there are so few points visible because of the configuration of the graph to not allow overlapping data points. The graph shows that as highway milage is increased, city milage is also increased. This makes sense because a car that is more efficient in general will be more efficient in both highway and city speeds.
```{r}
plot(mpg$hwy, mpg$cty)
```

4. Are cars in 2008 more fuel efficient than cars in 1999? Use graphs *and* summary statistics to answer this question. Explain your findings.

I took the average of the milage both in city and highway between the years 1999 and 2008 and compared. The highway milage is essentially equal, but there is a very slight advantage for vehicles in 1999 in city speeds.
```{r}
mpg99 = mpg[which(mpg$year == 1999), ]
mpg08 = mpg[which(mpg$year == 2008), ]

hwy99 = mean(mpg99$hwy)
cty99 = mean(mpg99$cty)

hwy08 = mean(mpg08$hwy)
cty08 = mean(mpg08$cty)

hwy99 = mean(mpg99$hwy)
cty99 = mean(mpg99$cty)
```

```{r}
hwy08
hwy99
cty08
cty99
```

I split the earlier plot into the 99 and 08 milages with highway on the x-axis and city milage on the y-axis. The graphs have a very similar shape and slope. However, there are a few outliers for 1999 where the city milage is near 35 and highway is near 45 that the 2008 data is lacking.
```{r}
plot(mpg99$hwy, mpg99$cty)
plot(mpg08$hwy, mpg08$cty)
```

Overall, looking at the averages and the graphs shows there is almost no difference in fuel efficiency between cars in 2008 and 1999.

Note: your submission is supposed to be fully reproducible, i.e. the TA and I will 'knit' your submission. Make sure that you don't use any file structure that depends on your computing environment.